### 2025-08-29 sections/4-results.tex
acknowleged unbalanced vs balanced section in discussion

In addition to the scaling experiments, we also explore how four different clustering strategies affect the performance of search.
These strategies are the Cartesian product of balanced vs. unbalanced clustering, and the presence vs. absence of depth-first reordering as described in Section~\ref{sec:methods:clustering:depth-first-reordering}.
To help with this analysis, we also added instrumentation to our implementation of CAKES to count the number of distance computations during search.
We note that this instrumentation significantly slows down the wall-clock time of search, so it would not be used in a real-world application.

Figure~\ref{fig:results:distance-counts}. shows these four comparisons (referred to as Ball, BalancedBall, PermutedBall, and, PermutedBalancedBall, where depth-first-reordering is called Permuted for brevity) for all three search algorithms on the Fashion-MNIST dataset.
Notably, performance is not only strictly worse for the balanced approaches, but the asymptotic behavior is significantly worse;
the algorithms (such as Depth-First Sieve) that seemed to exhibit constant-time behavior in Figure~\ref{fig:results:scaling-plots} ( which shows results using PermutedBall), no longer do so under a balanced clustering.
We note that on this dataset, depth-first reordering (the Permuted variants) seems to have little effect on throughput; depth-first reordering is primarily intended to improve space complexity.

>>>>

Beyond scaling behavior, we examine how clustering strategy affects search performance. We consider four strategies—the Cartesian product of balanced vs.\ unbalanced clustering and the presence vs.\ absence of depth-first reordering (Sec.~\ref{sec:methods:clustering:depth-first-reordering}). We instrument CAKES to count distance computations during search; this adds substantial overhead and would not be used in real-world applications.

Figure~\ref{fig:results:distance-counts} compares \textit{Ball}, \textit{BalancedBall}, \textit{PermutedBall}, and \textit{PermutedBalancedBall} (``Permuted'' denotes depth-first reordering) for all three algorithms on Fashion-MNIST. Balanced clustering is strictly worse in throughput and has poorer asymptotics: algorithms such as Depth-First Sieve that appear nearly constant-time in Fig.~\ref{fig:results:scaling-plots} (with \textit{PermutedBall}) no longer do so under balanced clustering. On this dataset, depth-first reordering has little effect on throughput; it is primarily intended to improve space complexity.
